Intentional release of native species undermines ecological stability

### 2022-10-05 rmd/figures.Rmd
Squashed commit of the following:
commit c274913bbbf475b2379bfd5ff5714134b1c4d743
Author: aterui <hanabi0111@gmail.com>
Date:   Wed Sep 14 12:52:06 2022 -0400

    edit: rmd misc

commit b83fa58b0ac121cf4f89fff102048b2ff07a3b1e
Author: aterui <hanabi0111@gmail.com>
Date:   Wed Aug 17 18:21:49 2022 -0400

    Update mcmc_trace_reg.pdf

commit 184921982330ea83d449b2fe44159659f9c52210
Author: aterui <hanabi0111@gmail.com>
Date:   Wed Aug 17 18:21:45 2022 -0400

    edit: misc

commit aaac7131a165551933b99ae948f54d710a673206
Author: aterui <hanabi0111@gmail.com>
Date:   Wed Aug 17 18:21:32 2022 -0

@@

```{r figure_1, fig.width=9, fig.height=11}

source(here::here("code/figure_theory_obs.R"))

print(g)

**Figure 1** Theory and empirical observations agree that intentional release destabilizes community dynamics. In A and B, panels and colors distinguish response variables and species groups. (A) Theoretical predictions. Dots represent individual simulation replicates, and lines and shades are loess curves fitted to simulated data and 95% confidence intervals. Parameters used in this simulation are: intrinsic growth rate of an enhanced species $r_1=$ `r unique(df0$r1)`, average interspecific competition $\bar{\alpha}=$ `r unique(df0$alpha)`, carrying capacity $K=$ `r unique(df0$k)`, environmental stochasticity $\sigma_{\epsilon}=$ `r unique(df0$sd_env)`, relative fitness of captive-bred individuals $f_R=$ `r unique(df0$phi)`. (B) Empirical evidence. Dots represent geometric means of site-level observations in 31 watersheds. Lines are the predicted values of the regression models, and line types correspond to the coefficient's posterior probabilities. See Tables S8-10 for full statistics. (C) Map of sampling sites (black dots) in Hokkaido, Japan. (D) Example of a protected watershed (Okushibetsu). (E) Masu salmon *Oncorhynchus masou masou*. Photo credit: Akira Terui.

@@

source(here::here("code/figure_coef.R"))

print(g_coef)

@@

**Figure 2** Posterior distributions for the standardized regression coefficients of intentional release. Y-axis represents different response variables grouped by colors distinguishing species groups (whole community, enhanced species [masu salmon], and unenhanced species).

>>>>

source(here::here("code/set_functions.R"))

@@

```{r figure1, fig.width=9, fig.height=11}

source(here::here("code/figure_layout.R"))

knitr::include_graphics(here::here("output/figure_theory_obs.pdf"))
```

**Figure 1** Theory and empirical observations agree that intentional release destabilizes community dynamics. In A and B, panels and colors distinguish response variables and species groups. (A) Theoretical predictions. Dots represent individual simulation replicates, and lines are the loess curves fitted to simulated data with their 95% confidence (dark shade) and 50% prediction intervals (light shade). Parameters used in this simulation are: intrinsic growth rate of an enhanced species $r_1=$ `r round(na.omit(unique(df0$r1)), 2)`, average interspecific competition $\bar{\alpha}=$ `r round(na.omit(unique(df0$alpha)), 2)`, carrying capacity $K=$ `r round(na.omit(unique(df0$k)), 2)`, environmental stochasticity $\sigma_{\epsilon}=$ `r round(na.omit(unique(df0$sd_env)), 2)`, relative fitness of captive-bred individuals $f_R=$ `r round(na.omit(unique(df0$phi)), 2)`. These parameter values are comparable to the estimates of the multi-species Ricker model. (B) Empirical evidence. Dots represent observations at 97 sites within 31 watersheds. Lines and shades are the predicted values and their 95% credible intervals of the regression models, and line types correspond to the coefficient's posterior probabilities. Effective release is a latent explanatory variable estimated in our hierarchical Baysian model. See Methods for details. Full statistics were reported in Tables S6-8. (C) Map of sampling sites (black dots) in Hokkaido, Japan. (D) Example of a protected watershed (Okushibetsu). (E) Masu salmon *Oncorhynchus masou masou*. Photo credit: Akira Terui.

@@

source(here::here("code/figure_obs_coef.R"))

knitr::include_graphics(here::here("output/figure_coef.pdf"))

@@

**Figure 2** Posterior distributions for the standardized regression coefficients of intentional release. Y-axis represents different response variables grouped by colors distinguishing species groups (whole community, enhanced species [masu salmon], and unenhanced species). Triangles indicate median estimates.

```{r figure3, fig.width=9, fig.height=11}

source(here::here("code/si_figure_species_int.R"))

knitr::include_graphics(here::here("output/figure_alpha_dist.pdf"))
```

**Figure 3** Estimated competition coefficients of the multi-species Ricker model. Dots represent a given pair of taxa, and box plots show median (thick vertical lines) and quantile (hinges). Whiskers show smallest (largest) observation less (greater) than or equal to upper hinge - 1.5 \* quantile (+ 1.5 \* quantile). The vertical line denotes the median. The density plot on the top shows the overall distribution of the competition coefficient. Colors distinguish taxa.

### 2022-10-05 rmd/manuscript.Rmd
edit: text

The past decades have witnessed efforts to unveil ecological risks associated with massive releases of captive-bred native species ("intentional release"). Yet, this method is still pervasive in conservation and natural resource management, assuming that the demographic benefits exceed the associated risks. However, such external inputs may disrupt the sensitive balance of species interactions that allows competing species to coexist, potentially compromising the long-term community stability. Here, we use theory and long-term data of stream fish communities to show that intentional release triggers community collapse with little demographic benefit to the enhanced species. Our theory predicted that intentional release intensifies interspecific competition, facilitating competitive exclusion of unenhanced species that otherwise stablely coexist. In parallel, the excessive input of captive-bred individuals suppressed the natural recruitment of the enhanced species via intensified within-species competition. Consequently, the ecological community with intentional release is predicted to show reduced community density with unstable temporal dynamics. Consistent with this prediction, stream fish communities showed greater temporal fluctuations and fewer species richness in rivers with the intensive release of hatchery salmon - a major fishery resource worldwide. Our findings alarm that the current overreliance on intentional release may accelerate global biodiversity loss, threatening the provisioning of ecosystem services.

@@

Human demands for natural resources are ever-increasing, such that active interventions are critical to the sustainable management of fisheries, forestry, and wildlife [@laikreCompromisingGeneticDiversity2010]. Releases of captive-bred native species ("intentional release") is a form of the efforts to enhance wild populations of diverse taxa [@snyderLimitationsCaptiveBreeding1996; @laikreCompromisingGeneticDiversity2010]. The magnitude of intentional release for commercial and recreational species is massive; typically, a single program releases tens of thousands to millions of plant, invertebrate (bivalves @haagNorthAmericanFreshwater2012, insects), and vertebrate individuals (fish, birds) into the wild annually (reviewed in [@laikreCompromisingGeneticDiversity2010; @kitadaEconomicEcologicalGenetic2018]). There is growing awareness that intentional release entails ecological risks, including the accumulation of deleterious alleles [@arakiGeneticEffectsCaptive2007], the spread of diseases and parasites in wild populations [@krkosekDecliningWildSalmon2007], and the intensification of intraspecific competition within the enhanced species. Yet, this method is still pervasive in natural resource management likely because of the apparent economic benefit of release programs. For example, Kitada @kitadaEconomicEcologicalGenetic2018 analyzed economic performance of release programs for 12 major fishery resources -- the gross monetary yield exceeded the seed-production cost in nine species, suggesting that release programs are economically profitable in general.

@@

Prior to our full-community analysis, we analyzed a two species community to understand the model behavior. The prior analysis revealed distinct parameter spaces predicting destabilizing, neutral, and stabilizing effects of intentional release on community dynamics. In general, intentional release had a neutral to destabilizing effect in a community with slow-growing species ($r_i < 1.5$). Meanwhile, for higher values of $r_i$ ($\ge 1.5$), intentional release stabilized community dynamics by dampening limit cycles or chaos. Carrying capacity and competition also played an important role in shifting the parameter spaces. According to this result, we chose `r nrow(list_param[[2]])` simulation scenarios, differing in the intrinsic growth of an enhanced species $r_1$, competition $\bar{\alpha}$, carrying capacity $K$, and relative fitness $f_R$ (see **Methods**). Under each parameter scenario, we ran `r attr(list_param[[2]], "n_sim")` time steps of `r max(list_sim[[2]]$n_rep)` independent communities (i.e., simulation replicates). Using the last `r unique(list_sim[[2]]$n_timestep)` time steps, we obtained the following summary statistics of the total community density $\sum_i^SN_i$ to examine the community-level response to intentional release: coefficient of variation (CV), number of species persist (defined as $N_i > 0.01$ for the entire simulation), temporal mean ($\mu$), and temporal SD ($\sigma$). We also calculated the temporal mean and SD for the enhanced ($N_1$) and unenhanced species ($\sum_{i,i \ne 1}^S N_i$) separately to infer underlying mechanisms.

When carrying capacity was small ($K=$ `r min(list_param[[2]]$k)`), our model predicted a destabilizing effect of intentional release on ecological communities as illustrated by increased CV with increasing numbers of releases (**Figure 1A**). This pattern stemmed mainly from the reduced mean of the total community density, and both enhanced and unenhanced species groups were responsible (**Figure 1A**). The enhanced species decreased because release induced the negative competitive effect exceeding the reproductive contribution of released individuals [@satakeStockingCaptivebredFish2012]. Meanwhile, interspecific competition reduced the unenhanced species at high levels of intentional release, resulting in fewer persisting species (**Figure 1A**). Combined, the total community density decreased more sharply than individual species groups (**Figure 1A**). The SDs showed a similar trend, but the relationship was flatter at the community level (**Figure 1A**). Since a CV is a ratio of an SD to a mean, the steeper decline of the mean community density led to the increased CV. These patterns were qualitatively similar across ecological contexts with the low carrying capacity (robust to the changes in $r_1$, $\bar{\alpha}$, and $f_R$; **Figures S1-S3**).

@@

Intentional release, however, had little or a stabilizing influence on community dynamics when carrying capacity was sufficiently large ($K=$ `r max(list_param[[2]]$k)`; **Figures S4-S6**). In particular, intentional release increased the enhanced species with a low population growth rate ($r_1=$ `r min(list_param[[2]]$r1)`; **Figure S4**), suggesting that weak intraspecific competition within the enhanced species (= $r_1 K^{-1}$) is needed for intentional release to be effective. Our result is reasonable because some of the best evidence for successful intentional release comes from re-introduction/translocation programs of extirpated species or population enhancement of long-lived endangered species. In the Sado island (Japan), for example, the endangered Crested Ibis *Nipponia nippon* showed exponential growth of the population since the initial reintroduction of ten captive-bred individuals in 2008. Similarly, translocations of Humpback chub, a federally-listed fish species, seem highly cost-effective in enhancing their tributary populations in the Little Colorado River. Two common features exist in these successful cases. First, the receiving ecosystem remains nearly-intact (i.e., high carrying capacity). Second, the number of releases in these conservation-oriented programs is rather small (several tens to hundreds) unlike release programs for commercial and recreational species (up to several tens of millions). These conditions meet the theoretical assumptions with which intentional release led to increased abundance of the enhanced species. However, the preponderance of unsuccessful reintroduction programs implies that these conditions are not always sufficed. In several examples, excessive numbers of individuals are released into degraded ecosystems with absolute failures of population enhancement or establishment. Our analysis suggests that pre-release examination and restoration of environmental capacity may be a key to successful release programs with minimal impacts on other community members.

@@

As predicted, stream fish communities showed greater temporal fluctuations (higher CV) in watersheds with intensive release (**Figure 1B**). The effect was striking in its magnitude, almost doubling the CV at the highest release level. Our analysis strongly supported the positive relationship between the CV and the number of releases (**Figure 2**), in which the probability of the regression coefficient being positive was XXX (**Table S8**). This pattern was associated with the reduced long-term average of the total community density and fewer species richness (**Figures 1B and 2**), and both enhanced (masu salmon) and unenhanced fish groups contributed to this trend (**Figures 1B and 2;** see **Tables S8-S10** for full statistics). In the meantime, the SDs had vague relationships with intentional release (**Figures 1B and 2**).

To gain better insights into the driving mechanisms, we fitted a multi-species Ricker model to a subset of our time-series data at nine sites with few missing observations and no intentional release (**Methods**); thus, the parameter estimates of this model is directly comparable to those used in the theoretical Ricker model. The estimated competition coefficients $\alpha_{ij}$ ranged `r round(min(df_pd$alpha_prime), 2)` -- `r round(max(df_pd$alpha_prime), 2)` with the median of `r round(median(df_pd$alpha_prime), 2)` (Figure 3), suggesting that competition is a likely mechanism as our theory assumes. Our statistical inference is supported by previous observational studies. For example, interference competition between masu salmon and white-spotted charr is well studied. Their diets are innately overlapped (e.g., drift terrestrial insects), but they choose different prey items when they co-occur [@miyasakaFoodHabitDivergence2003]. Exploitative competition between masu salmon and benthic fish (e.g., sculpins) is also likely because subordinate masu salmon in the dominance hierarchy consumes benthic macroinvertebrates [@satoStagespecificFunctionalResponses2014], which are dominant prey items for benthic fish species. Importantly, hatchery masu salmon may intensify intra- and interspecific competition in the wild [@hasegawaWildMasuSalmon2018; @sahashiEvaluationContributionHatchery2015; @hasegawaPulsedSuppliesSmall2022]. Hatchery salmons are larger and more aggressive than wild individuals, increasing the likelihood of intense intra- and interspecific competition [@weberInteractionsHatcheryWild2003]. Indeed, field and experimental studies confirmed that hatchery masu salmon competed with wild masu salmon and other stream fishes during their freshwater life stage [@hasegawaWildMasuSalmon2018; @sahashiEvaluationContributionHatchery2015; @hasegawaPulsedSuppliesSmall2022].

The reduced fitness of hatchery masu salmon may also play a role in the observed response. Like other salmonids, ocean-migrating adults of hatchery masu salmon show lower return rates to the spawning river [@sahashiWildGenesBoostinpress]. In addition, the aggressive behavior of hatchery salmons may make them vulnerable to predation [@weberInteractionsHatcheryWild2003]. The fitness disadvantage may add to negative density dependence to influence community stability.

Importantly, socioeconomic factors control the number of releases (e.g., budget allocation) regardless of the current condition of recipient communities. As such, released fish are probably "excessive" and may cause resource competition that would otherwise not exist.

We cannot exclude the possibility that spurious correlations drove our results. However, the protected watersheds preserve nearly intact landscapes (**Figure S10**) with strict regulations of human activities (e.g., exploitation, construction of in-stream structures); therefore, it is difficult to envision that unmeasured human influences caused the observed relationships. Further, we have statistically controlled natural variation in environmental factors (see **Methods** and **Tables S8-S10**). This unique setup may have helped uncover the qualitative agreement between theory and empirical patterns.

Despite the significant attention to the fate of captive-bred individuals [@arakiHatcheryStockingHelp2010; @snyderLimitationsCaptiveBreeding1996], current schemes rarely consider the self-regulation process of biodiversity. Our results suggest that the ignorance of this critical process may erode the long-term persistence of the recipient community, likely impacting the provisioning of ecosystem services [@cardinaleBiodiversityLossIts2012]. While our empirical example is limited to stream fish communities, we anticipate that this phenomenon is pervasive in nature because the destabilizing effect emerged across diverse simulated scenarios.

We should prioritize habitat conservation with a broader scope of ecosystem management for the sustainability of natural resources. Such efforts are particularly important in the Anthropocene because habitat degradation may exacerbate the undesired influence of intentional release. Protected areas and environmental restoration are promising tools to conserve biodiversity, and a smart spatial design is a key to achieving successful conservation. For example, coordinated placement of conservation sites considering spatial biodiversity patterns is crucial in improving the ecological outcomes [@teruiEmergentDualScaling2021; @koningNetworkGrassrootsReserves2020; @fontouraProtectingConnectivityPromotes2022]. Governance may also play a central role in enforcing environmental legislation, potentially determining the effectiveness of conservation investment [@amanoSuccessfulConservationGlobal2018]. These considerable potentials indicate that viable management options exist before blindly accepting intentional release. Without a comprehensive framework that appreciates the ecological integrity of natural communities, the intentional release will never be effective but impairs biodiversity.

>>>>

Massive releases of captive-bred native species ("intentional release") is a pervasive method to enhance wild populations of commercial and recreational species. However, such external inputs may disrupt the sensitive balance of species interactions that allows competing species to coexist, potentially compromising the long-term community stability. Here, we use theory and long-term data of stream fish communities to show that intentional release triggers community collapse with little demographic benefit to the enhanced species. Our theory predicted that intentional release intensifies interspecific competition, facilitating competitive exclusion of unenhanced species that otherwise stablely coexist. In parallel, the excessive input of captive-bred individuals suppressed the natural recruitment of the enhanced species via intensified within-species competition. Consequently, the ecological community with intentional release is predicted to show reduced community density with unstable temporal dynamics. Consistent with this prediction, stream fish communities showed greater temporal fluctuations and fewer species richness in rivers with the intensive release of hatchery salmon - a major fishery resource worldwide. Our findings alarm that the current overreliance on intentional release may accelerate global biodiversity loss, threatening the provisioning of ecosystem services.

@@

Human demands for natural resources are ever-increasing, such that active interventions are critical to the sustainable management of fisheries, forestry, and wildlife [@laikreCompromisingGeneticDiversity2010]. Releases of captive-bred native species ("intentional release") is a form of the efforts to enhance wild populations of diverse taxa [@snyderLimitationsCaptiveBreeding1996; @laikreCompromisingGeneticDiversity2010]. The magnitude of intentional release for commercial and recreational species is massive; a single program releases tens of thousands to millions of plant [@laikreReleaseAlienPopulations2006], invertebrate (bivalves, insects) [@ingsBumblebeesHumblePollinators2005; @haagNorthAmericanFreshwater2012], and vertebrate individuals (fish, birds) [@villanuaSanitaryRisksRedlegged2008; @kitadaEconomicEcologicalGenetic2018] into the wild annually (reviewed in [@laikreCompromisingGeneticDiversity2010; @kitadaEconomicEcologicalGenetic2018]). There is growing awareness that intentional release entails ecological risks, including the accumulation of deleterious alleles [@arakiGeneticEffectsCaptive2007], the spread of diseases and parasites in wild populations [@krkosekDecliningWildSalmon2007], and the intensification of intraspecific competition within the enhanced species [@kitadaLessonsLearnedJapanese2006]. Yet, this method is still pervasive in natural resource management likely because of the apparent economic benefit of release programs. For example, Kitada @kitadaEconomicEcologicalGenetic2018 analyzed economic performance of release programs for 12 major fishery resources -- the gross monetary yield exceeded the seed-production cost in nine species, suggesting that release programs are economically profitable in general.

@@

Prior to our full-community analysis, we analyzed a two species community to understand the model behavior. The prior analysis revealed distinct parameter spaces predicting destabilizing, neutral, and stabilizing effects of intentional release on community dynamics (**Figures S1 and S2**). In general, intentional release had a neutral to destabilizing effect in a community with slow-growing species ($r_i < 1.5$). Meanwhile, for higher values of $r_i$ ($\ge 1.5$), intentional release stabilized community dynamics by dampening limit cycles or chaos. Carrying capacity and competition also played an important role in shifting the parameter spaces. According to this result, we chose `r nrow(list_param[[2]])` simulation scenarios, differing in the intrinsic growth of an enhanced species $r_1$, competition $\bar{\alpha}$, carrying capacity $K$, and relative fitness $f_R$ (see **Methods**). Under each parameter scenario, we ran `r attr(list_param[[2]], "n_sim")` time steps of `r max(list_sim[[2]]$n_rep)` independent communities (i.e., simulation replicates). Using the last `r unique(list_sim[[2]]$n_timestep)` time steps, we obtained the following summary statistics of the total community density $\sum_i^SN_i$ to examine the community-level response to intentional release: coefficient of variation (CV), number of species persist (defined as $N_i > 0.01$ for the entire simulation), temporal mean ($\mu$), and temporal SD ($\sigma$). We also calculated the temporal mean and SD for the enhanced ($N_1$) and unenhanced species ($\sum_{i,i \ne 1}^S N_i$) separately to infer underlying mechanisms.

When carrying capacity was small ($K=$ `r min(list_param[[2]]$k)`), our model predicted a destabilizing effect of intentional release on ecological communities as illustrated by increased CV with increasing numbers of releases (**Figure 1A**). This pattern stemmed mainly from the reduced mean of the total community density, and both enhanced and unenhanced species groups were responsible (**Figure 1A**). The enhanced species decreased because release induced the negative competitive effect exceeding the reproductive contribution of released individuals [@satakeStockingCaptivebredFish2012]. Meanwhile, interspecific competition reduced the unenhanced species, resulting in fewer persisting species (**Figure 1A**). Combined, the total community density decreased more sharply than individual species groups (**Figure 1A**). The SDs showed a similar trend, but the relationship was flatter at the community level (**Figure 1A**). Since a CV is a ratio of an SD to a mean, the steeper decline of the mean community density led to the increased CV. These patterns were qualitatively similar across ecological contexts with the low carrying capacity (robust to the changes in $r_1$, $\bar{\alpha}$, and $f_R$; **Figures S3-S6**).

@@

Intentional release, however, had little or a stabilizing influence on community dynamics when carrying capacity was sufficiently large ($K=$ `r max(list_param[[2]]$k)`; **Figures S7-S10**). In particular, intentional release increased the enhanced species with a low population growth rate ($r_1=$ `r min(list_param[[2]]$r1)`; **Figure S7**), suggesting that weak intraspecific competition within the enhanced species (= $r_1 K^{-1}$) is needed for intentional release to be effective. Our result is reasonable because some of the best evidence for successful intentional release comes from re-introduction/translocation programs of extirpated species or population enhancement of long-lived endangered species [@deguchiTranslocationHandrearingResult2017; @jonesRestorationMauritiusKestrel1995; @okahisaEvaluationOngoingCrested2022; @yackulicAssessingPopulationImpacts2021]. In the Sado island (Japan), for example, the endangered Crested Ibis *Nipponia nippon* showed exponential growth of the population since the initial reintroduction of ten captive-bred individuals in 2008 [@okahisaEvaluationOngoingCrested2022]. Similarly, translocations of Humpback chub, a federally-listed fish species in the US, seem highly cost-effective in enhancing their tributary populations in the Little Colorado River [@yackulicAssessingPopulationImpacts2021]. Two common features exist in these successful cases. First, the receiving ecosystem supports suitable habitats (i.e., high carrying capacity). Second, the number of releases in these conservation-oriented programs is rather small (several tens to hundreds) [@fischerAssessmentPublishedResults2000] in comparison to release programs for commercial and recreational species (up to several tens of millions). These conditions meet the theoretical assumptions with which intentional release led to increased abundance of the enhanced species. However, the preponderance of unsuccessful reintroduction programs implies that these conditions are not always sufficed [@fischerAssessmentPublishedResults2000; @haagNorthAmericanFreshwater2012; @strayerEssayMakingMost2019]. Indeed, non-negligible numbers of projects seem to release captive-bred individuals into the wild without knowing whether the release habitat is suitable for the focal species [@fischerAssessmentPublishedResults2000; @haagNorthAmericanFreshwater2012]. Our analysis suggests that pre-release examination and restoration of environmental capacity may be a key to successful release programs with minimal impacts on other community members.

@@

As predicted, stream fish communities showed greater temporal fluctuations (higher CV) in watersheds with intensive release (**Figure 1B**). The effect was striking in its magnitude, almost doubling the CV at the highest release level. Our analysis strongly supported the positive relationship between the CV and the number of releases (**Figure 2**), in which the probability of the regression coefficient being positive was `r op(df_beta %>% filter(response == "cv") %>% pull(prob), 2)` (**Table S8**). This pattern was associated with the reduced long-term average of the total community density and fewer taxonomic richness (**Figures 1B and 2**), and both enhanced (masu salmon) and unenhanced fish groups contributed to this trend (**Figures 1B and 2;** see **Tables S6-S8** for full statistics). In the meantime, the SDs had vague relationships with intentional release (**Figures 1B and 2**).

To gain better insights into the driving mechanisms, we fitted a multi-species Ricker model to a subset of our time-series data at nine sites with few missing observations and no intentional release (**Methods**). The parameter estimates of this model are thus directly comparable to those used in the theoretical Ricker model. The estimated competition coefficients $\alpha_{ij}$ ranged `r op(min(df_pd$alpha_prime), 2)` -- `r op(max(df_pd$alpha_prime), 2)` with the median of `r op(median(df_pd$alpha_prime), 2)` (**Figure 3**), suggesting that competition is a likely mechanism as our theory assumes (see **Figure S17** for all pairwise estimates). Our statistical inference is supported by previous studies. For example, interference competition between masu salmon and other salmonid species is well documented [@miyasakaFoodHabitDivergence2003; @hasegawaComparisonCompetitiveAbility2004a]. Diets of salmonid species are innately overlapped (e.g., drift terrestrial insects), such that they interfere with each other for preferable foraging microhabitats [@hasegawaComparisonCompetitiveAbility2004a]. Exploitative competition between masu salmon and benthic fish (e.g., sculpins and stone loaches in our study system) is also likely because subordinate masu salmon in the dominance hierarchy consumes benthic macroinvertebrates [@satoStagespecificFunctionalResponses2014], which are dominant prey items for many benthic fish species [@natsumedaEvaluationEcologicalFeatures2010; @natsumedaDietFoodSelectivity2012]. Importantly, hatchery masu salmon may intensify intra- and interspecific competition in the wild [@hasegawaWildMasuSalmon2018; @sahashiEvaluationContributionHatchery2015; @hasegawaPulsedSuppliesSmall2022]. Hatchery salmons are larger and more aggressive than wild individuals at a comparable life stage, increasing the likelihood of intense intra- and interspecific competition [@weberInteractionsHatcheryWild2003]. Indeed, field and experimental studies confirmed that hatchery masu salmon competed with wild masu salmon and other stream fishes during their freshwater life stage [@hasegawaWildMasuSalmon2018; @sahashiEvaluationContributionHatchery2015; @hasegawaPulsedSuppliesSmall2022].

The reduced fitness of hatchery masu salmon may also play a role in the observed response. Like other salmonids, masu salmon exhibits a life history of partial migration [@moritaLatitudinalVariationGrowth2010; @sahashiMigrationCostsDrive2013]. In our study region, almost all females and the majority of males migrate to the ocean and spend approximately one year before spawning, although some males stay in freshwater for their entire life [@moritaLatitudinalVariationGrowth2010; @moritaTemperaturedependentVariationAlternative2014]. It has been shown that ocean-migrating adults of hatchery masu salmon show lower return rates to the spawning river [@sahashiWildGenesBoostinpress]. In addition, the aggressive behavior of hatchery salmons may make them vulnerable to predation [@weberInteractionsHatcheryWild2003]. The fitness disadvantage may add to negative density dependence to influence community stability.

Importantly, socioeconomic factors control the number of releases (e.g., budget allocation) regardless of the current condition of recipient communities. As such, released fish are probably "excessive" and may cause resource competition that would otherwise not exist. XXX -

It is noteworthy that the magnitude of the release program at the protected watersheds (annual average maximum = `r op(max(df_stock_mu$mean_stock), 2)` million fish) is comparable or even smaller when compared with those for other commercial and recreational species [@laikreCompromisingGeneticDiversity2010; @kitadaEconomicEcologicalGenetic2018]. For example, 0.5 -- 1.2 million juveniles of Black sea bream are released annually in the Kagoshima bay, Japan, where substantial declines of genetic diversity and wild individuals have been reported [@kitadaRigorousMonitoringLargescale2019]. Importantly, a meta-analysis focusing on fishery resources revealed that only a few of these massive release programs led to increased abundance of the enhanced species [@arakiHatcheryStockingHelp2010]. Hence, the patterns observed in the protected watersheds are likely to be general, although none of those studies quantified the community-level impact of massive releases. Future studies addressing this knowledge gap is therefore highly valuable.

The near-intact environment of the protected watersheds with strict regulations of human activities (**Figure S10**) may have helped uncover the qualitative agreement between theory and empirical patterns. For example, any form of exploitation, including angling, is strictly prohibited for any fish species in the protected watersheds. In addition, an exclusive permission is required to physically alter in-stream and riparian habitats (e.g., construction of river-crossing structures). Therefore, unmeasured human influences should have minimal influences on the observed relationships. Further, we have statistically controlled natural variation in environmental factors (see **Methods**). Nevertheless, our results must be viewed with caution: we cannot exclude the possibility of spurious correlations due to the inherent nature of correlative analysis and field research. Since an experimental approach is nearly impossible at this spatial scale, the application of novel causal inferences (e.g., convergent cross mapping) [@sugiharaDetectingCausalityComplex2012; @ushioFluctuatingInteractionNetwork2018] may be a potential alternative to confirm the causal relationship between intentional release and community dynamics if sufficient time-series data are available.

Despite the significant attention to the fate of captive-bred individuals [@arakiHatcheryStockingHelp2010; @snyderLimitationsCaptiveBreeding1996], current schemes rarely consider the self-regulation process of biodiversity. Our results suggest that the ignorance of this critical process may erode the long-term persistence of the recipient community, likely impacting the stable delivery of ecosystem services [@cardinaleBiodiversityLossIts2012]. While more sophisticated analysis is required to provide detailed guidance on how to improve release programs, it is clear from our analysis that habitat conservation should be prioritized with a broader scope of ecosystem management for the sustainability of natural resources. Such efforts are particularly important in the Anthropocene because habitat degradation may exacerbate the undesired influence of intentional release by compromising the environmental capacity. Protected areas and environmental restoration are promising tools to conserve biodiversity, and a smart spatial design is a key to achieving successful conservation. For example, coordinated placement of conservation sites considering spatial biodiversity patterns is crucial in improving the ecological outcomes [@teruiEmergentDualScaling2021; @koningNetworkGrassrootsReserves2020; @fontouraProtectingConnectivityPromotes2022]. Governance may also play a central role in enforcing environmental legislation, potentially determining the effectiveness of conservation investment [@amanoSuccessfulConservationGlobal2018]. These considerable potentials indicate that viable management options exist before blindly accepting intentional release. Without a comprehensive framework that appreciates the ecological integrity of natural communities, the intentional release will never be effective but impairs biodiversity.
